Multivariate Skew-Normal Mixture Model for Infant Development Clustering

### 2019-05-29 manuscript/manuscript.tex
references
poster

\title[]{A multivariate skew-normal finite mixture model for analysis of infant development trajectories}

@@

@@

@@

@@

\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $$ t $$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.

>>>>

\title[]{Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories}

@@

\subsection{Existing Approaches}

Mixtures of multivariate non-symmetric distributions such as the multivariate skew-normal (MSN) distribution allow for the nuances of the marginal density to be captured with a more parsimonious set of mixture components. Mixtures of MSN distributions has been dealt with previously in a Bayesian context (Fr\"{u}hwirth-Schnatter \& Pyne, 2010), however in these models, focus lies primary on marginal density estimation and inference on the mixture components (i.e. clusters) is not discussed. More recently, the mixtures of skew-$t$ factor analysis (MSTFA) model has been proposed be applied in settings where cluster-specific inference is of primary interest. However, thid

@@

\bibitem{ } Arellano‐Valle RB, Azzalini A. On the unification of families of skew‐normal distributions. \textit{Scandinavian Journal of Statistics}. 2006 Sep;33(3):561-74.

\bibitem{ } Azzalini A. A class of distributions which includes the normal ones. \textit{Scandinavian journal of statistics}. 1985 Jan 1:171-8.

\bibitem{ } Azzalini, A. and Dalla Valle, A. (1996). The multivariate skew normal distribution. Biometrika 83, 715–726.

@@

\bibitem{ } Lee SX, Mclachlan GJ. On mixtures of skew normal and skew $t$-distributions. \textit{Advances in Data Analysis and Classification}. 2013 Sep 1;7(3):241-66.

@@

### 2019-05-29 poster/poster.tex
manuscript format
outline
poster

\begin{block}{Motivation}
\begin{figure}
\includegraphics[width=0.9\linewidth]{bayley_dens.jpg}
\caption{Density of Bayley composite scores for Nurture infants at 3, 6, 9 and 12 months of age.}
\end{figure}
\end{block}

@@

$$P(Z_i = l|\w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$

@@

\begin{figure}
\includegraphics[height=0.9\linewidth,width=1\linewidth]{beta_trace.pdf}
\caption{Trace plots of regression parameter estimates in simulation}
\end{figure}

@@

@@

@@

\tiny \textbf{Funding:} \textit{This work is supported by a grant from the NIH (\textbf{R01DK094841})}.

>>>>

\begin{block}{Motivation}
\begin{figure}
\includegraphics[width=0.9\linewidth]{bayley_dens.jpg}
\caption{Residuals of repeated measures model of Bayley scores for infants at 3, 6, 9 and 12 months of age.}
\end{figure}
\end{block}

@@

$$P(Z_i = l|w_i) = \pi_{il} = \frac{e^{w_i^T \delta_l}}{\sum_{r = 1}^h e^{w_i^T \delta_r}}$$

@@

@@

\begin{table}[]
\caption{\label{tab:unnamed-chunk-4}Estimated effects of sex and race on odds of cluster membership relative to the reference cluster.}
\centering
\begin{tabular}{lll}
\hline
Variable & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 2\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Class 3\\ $\hat{\text{OR}}$ (95\% CrI)\end{tabular}} \\ \hline
Intercept & 1.28 (1.01, 1.73) & 1.16 (0.87, 1.71) \\
Sex (Female) & 0.69 (0.40, 0.96) & 0.77 (0.48, 1.05) \\
Race (White) & 0.90 (0.57, 1.32) & 0.83 (0.59, 0.29) \\ \hline
\end{tabular}
\end{table}

In Table 2, we demonstrate the capability of estimating the effect of covariates (race and sex here) on posterior probability of cluster membership. We use class 1 as the reference class.

@@

Above is a plot of the mean Bayley score for each of the fitted clusters. We observe similar development patterns but different baseline development between clusters 1 and 2, and a qualitatively different development pattern in class 3 than in the classes 1 and 2.

@@

\tiny \textbf{Funding:} \textit{This work is supported by a grant from the NIH (\textbf{R01DK094841}) and a grand from the NLM (\textbf{1R21LM012866-01})}.
